---
title: "conda"
author: "AY"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## conda - how to

The first step is to create a conda environment, containing the necessary tools for the required analysis. In this workflow, we use the `snakemake` package to help running the workflow, as well as some mappers. 
To learn more about the conda environment and its advantages, take a look at the [conda web site](https://conda.io/en/latest/).

Environments can be created with the help of a file, e.g. `environment.yaml` using the following command

```{r, engine='bash', eval=FALSE}
conda env create --name Project0815 --file environment.yaml
```
This file would look like that:

```
name: Project0815
channels:
   - bioconda
   - conda-forge
   - r
   - bioinfo
dependencies:
   - snakemake-minimal =5.2.4
   - python =3.6
   - star =2.7.2c
   - samtools =1.9
   - bowtie2 =2.3.5
   - bwa =0.7.17
```

The first row of the file contains the name of the project, i.e. the name of the conda environment created. Next are the `channels` listed, from which the conda environment should look for the tools needed for the analysis.

the last part of the environment file is a list of the needed `dependencies`. Each one of them has its version attached to it. This can be very helpful, when a data set needs to be re-analyzed to check for problems or conflicts with older/newer results or for debugging purposes.
In this workflow we will use the `snakemake` package as well as `python` version 3.6. we also install here the mappers `star`, `bwa` and `bowtie2`. WE also need the `samtool` package for the manipulation of the bam files (sorting, filtering, etc.). 

After this was done *one must activate the environment* using the command

```{r, engine='bash', eval=FALSE}
source activate Project0815
```

If one look at the command line one can now see this

```
(Project0815) USER@SERVERNAME:Project0815$ ..
```

The brackets on the left side must contains the name of the environment. 

If, after the environment was created, new tools must be added, they can just be added to the list of tools in the `environment.yaml` file and than the conda environment must be updated, using the command

```{r, engine='bash', eval=FALSE}	
conda env update -f environment.yaml 
```

The conda environment should be created only once, but with each login one must activate it, so that the snakemake script can be run from the correct paths.










